---
title: "Data Import"
output: github_document
---
# Lecture 4 code

```{r setup} 
# above: name the code chunk

library(tidyverse)
```


## Read in some data

1. Read Data: start from **project folder**, then to go to **data** folder, then open the data file.
```{r}
litters_df = read_csv("./data/FAS_litters.csv")  # ".": start from here (project folder), "/": go to

#Always followup step: 
litters_df = janitor::clean_names(litters_df) #load "janitor" package, then use "clean_name()" function
```

2. Always follow up by janitor::clean_names():
janitor::clean_names: clean up variable names after importing data. Doing so will take whatever the column names are and convert them to lower snake case.

## Take a look at the data

1. Printing in the console.
```{r}
litters_df
```
2. 
```{r}
head(litters_df) #show first 6 rows of the dataframe
tail(litters_df) #show the 6 rows of the dataframe
```
3. 
```{r}
skimr::skim(litters_df) 
#tells missing values, min, max, histograms, etc.

```
4. use view(datafram) function in the _console_ to open the file in R studio.

 - "character (chr)": the column that mix characters and numbers.
 - "double": a particular kind of _number_ in computer programming.
 
## options to read_csv
```{r}
litters_df = read_csv("./data/FAS_litters.csv", skip = 10, col_names = FALSE) #skip first 10 rows, and don't show the column names
```

## Other fiel formats

1. read in an excel file.
```{r}
library(readxl)
mlb_df = read_excel("./data/mlb11.xlsx")

#Read the excel file from A1 to E7
mlb_df = read_excel("./data/mlb11.xlsx", range = "A1:F7")

```

2. Read in an SAS file.
```{r}
library(haven)
pulse_df = read_sas("./data/public_pulse_data.sas7bdat")
```

## Comparison with Base R

`read.csv` vs `read_csv`? Never ever use `read.csv` !!!
- `read.csv` is the most racent version of R, which is easier to manage the data


