---
title: "Data manipulation"
output: github_document
---
##Lec6 code

1. dplyr package: manage data, included in library(tidyverse).
1)functions in dplyr will take a tibble as an input and return a tibble as a result

```{r message=FALSE}
library(tidyverse)
options(tibble.print_min = 5) #print only three lines of each tibble by default.
```

### import some data
```{r }
litters_df = read.csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```
 - 'janitor::clean_names()' always comes after import dataframe.  
 
### _`select`_ : operate on columns

1. Choose some columns and not others
```{r eval=FALSE}
select(litters_df, group, litter_number)
select(litters_df, group, gd0_weight, gd18_weight)
```


```{r eval=FALSE}
# select every column between gd0_weight and gd_of_birth `:` : 
select(litters_df, group, gd0_weight:gd_of_birth) 

#get rid of the column: litter_number, and keep everything others:
select(litters_df, -litter_number) 

```

2. Renaming columns
```{r eval=FALSE}
#select the `group` and `litter_number` columns; and rename `group` column to 'GROUP' and `litter_number` to `LITTer_NUmbe`
select(litters_df, GROUP = group, LITTer_NUmber = litter_number) 

#or keep every columns, just rename the specific column.
rename(litters_df, GROUP = group)
```

3. select helpers
```{r eval=FALSE}
#select everything which names are started by "pups"
select(litters_df, group, starts_with("pups"))
```

```{r eval=FALSE}
#ends_with()
select(litters_df, group, ends_with("pups"))
#contains()
select(litters_df, group, contains("pups"))
```

```{r eval=FALSE}
#select `litter_number` to the beginning of the dataframe and keep everything else
select(litters_df, litter_number, everything()) 
# or
relocate(litters_df, litter_number) #relocate `litter_number` to the beginning of the dataframe and keep everything else
```

4. `select()` vs. `pull()`
```{r eval=FALSE}
#select vectors and export a dataframe 
select(litters_df, group) 

#select vectors and store them into one column.
pull(litters_df, group) 
```


### _`filter`_: operate on rows   
 - Need to give conditions: TRUE/FALSE
1. One conditions: 
```{r eval=FALSE}
# keep everything else columns, but where gd_of_birth = 20, make sure two "=="
filter(litters_df, gd_of_birth == 20) 

# keep everything else columns, but where gd0_weight <= 34
filter(litters_df, gd0_weight <= 34)

# keep everything else columns, but where pup_survive not equal to 4
filter(litters_df, !(pups_survive == 4)) 
#or
filter(litters_df, pups_survive != 4) 

# keep everything else columns, but where groups are not Con7
filter(litters_df, !(group == "Con7")) 
```

2. follow two conditions:
```{r eval=FALSE}
#two conditions need to be follow
filter(litters_df, gd_of_birth == 20, group == "Con7") 
#two conditions need to be follow; "," and "&" has the same functions
filter(litters_df, gd_of_birth == 20 & group == "Con7") 
```

3. either one of the conditions will be followed
```{r eval=FALSE}
# the rows that either "gd_of_birth == 20" or "group =="Con7"" will be selected
filter(litters_df, gd_of_birth == 20 | group == "Con7") 

# %in%: the values in either one row or the other.
filter(litters_df, group %in% c("Con7", "Con8")) #the rows that groups in either "Con7" or "Con8", will be selected.

```

4. drop missing values   
```{r eval=FALSE}
# remove any row with a missing value
drop_na(litters_df)

#will remove rows for which `wt_increase` is missing.
drop_na(litters_df, gd0_weight) 
```

### Mutate
1. Modify the existing variables or create new variables:   
 - Add or change the columns
```{r eval=FALSE}
mutate(litters_df, 
       weight_change = gd18_weight - gd0_weight, # add a new column equals to gd18_weight - gd0_weight
       group = str_to_lower(group)) # Modify `group` variable and change everything in "group" column to lowercase 
```

### _`arrage()`_
1. rearrange the data: 
 - arrange by default is ascending; 
 - change to descending by using: desc()
```{r eval=FALSE}
arrange(litters_df, gd0_weight)
arrange(litters_df, desc(gd0_weight)) 
```


### Pipes
1. ` %>%`: "and then"   
 - The result of one function gets passed as the first argument to the next one by default.
```{r eval=FALSE}
#NOT using `%>%`: 
litters_data_row = read_csv("./data/FAS_litters.csv")
litters_clean_name = janitor::clean_names(litters_data_row)
litters_data_select = select(litters_clean_name, -pups_survive)
litters_mutated = mutate(litters_data_select, wt_gain = gd18_weight - gd0_weight)
litters_without_missing = drop_na(litters_mutated, gd0_weight)

#using PIPE operate: `%>%`
litters_df = 
  read_csv("./data/FAS_litters.csv") %>%
  janitor::clean_names() %>% 
  select(-pups_survive) %>% #no need to repeat dataset name
  mutate(wt_gain = gd18_weight - gd0_weight) %>% 
  drop_na(gd0_weight)
```

