---
title: "Data Import"
output: github_document
---

# Lecture 5

```{r setup} 
# name the code chunk as "setup"
library(tidyverse)
```


## Read in some data

1. Read Data: start from **project folder**, then to go to **data** folder, then open the data file.   
 - Hence, make sure to save the data files in the same directory or sub-directory of the project.
 
```{r}
litters_df = read_csv("./data/FAS_litters.csv")  
       # ".": start from here (project folder), "/": go to

#Always followup step: 
litters_df = janitor::clean_names(litters_df) #load "janitor" package, then use "clean_name()" function
```
 -  ~	Home directory   
    .	Current working directory   
    ..	One directory up from current working directory   
    ../..	Two directories up from current working directory   

2. REMEMBER: Always follow up by janitor::clean_names():
1) janitor::clean_names: clean up variable names after importing data. Doing so will take whatever the column names are and convert them to lower snake cases.
2) The `package::function` syntax: use a function from a package without loading the whole library.

## Take a look at the data

1. Printing in the console.
```{r}
litters_df
```

2. 
```{r}
head(litters_df) #show first 6 rows of the dataframe
tail(litters_df) #show the 6 rows of the dataframe
```

3. 
```{r}
skimr::skim(litters_df) 
#tells missing values, min, max, histograms, etc.
```

4. use `view(data frame)` function in the _console_ to open the file in R studio.   
1) "character (chr)": the column that mix characters and numbers.   
2) "double": a particular kind of _number_ in computer programming.   
 
## options to read_csv
```{r}
litters_df = read_csv("./data/FAS_litters.csv", skip = 10, col_names = FALSE) 
#skip first 10 rows, and don't show the column names
```

```{r}
litters_data = read_csv(file = "./data/FAS_litters.csv",
  col_types = "ccddiiii"
)
# read the file and make sure the column are in `c`: character, `d`: double, `i`: integer
```

```{r}
mean(pull(litters_data, `GD18 weight`), na.rm = TRUE ) 
#na: missing values
#exclude the missing valuse and count the mean.
```

## Other file formats

1. read in an excel file.
```{r}
library(readxl)
mlb_df = read_excel("./data/mlb11.xlsx")

#Read the excel file from A1 to E7
mlb_df = read_excel("./data/mlb11.xlsx", range = "A1:F7")
```

2. Read in an SAS file.
```{r}
library(haven)
pulse_df = read_sas("./data/public_pulse_data.sas7bdat")
```

## Comparison with Base R

`read.csv` vs `read_csv`? Never ever use `read.csv` !!!
- `read.csv` is the most recent version of R, which is easier to manage the data

## Exporting data as csv file

Export the mlb sub-table
```{r}
write_csv(mlb_df, "./data/mlb_subtable.csv")

#`write_` function is in `library(readr)`
# export `mlb_df` dataframe with the name: mlb_subtable.csv
```

