---
title: "tidy_data"
author: "Yuxuan Chen"
output: github_document
---

## Lecture 7
1. Rules for tidy data
 - Columns are variables; rows are observations   
 - Combine the columns / separate datasets to rows as much as possible
 
```{r message=FALSE}
library(tidyverse)
options(tibble.print_min = 5)
```

### `pivot_longer()`
1. read data:
```{r}
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%  #since this dataset is in the sas file, so use haven pacakge and `read_sas` funcion
  janitor::clean_names()
pulse_data
```

2. Let's try pivot   
 - combine all variables (columns) to each id: (wide format to long format)
 - relocate dataset to let `id` be first column, and `visit` be the second column
 -extend (add a mutate)
```{r}
pulse_data_tidy = 
  pulse_data %>%
  pivot_longer(
    bdi_score_bl:bdi_score_12m, #specify the range of variables(columns) I am interested in
    names_to = "visit", #put these variables(columns) into a new variable(column) named "visit"
    names_prefix = "bdi_score_",  #exclude the `bdi_score_` prefix in each case
    values_to = "bdi") %>% #put the values from these variables(columns) into a new variable(column) named "bdi"
  relocate(id, visit) %>% 
  mutate(visit = replace(visit, visit == "bl", "00 m"))    #change "bl" in "visit" column to "00m"; `recode()` change observations
```

### `pivot_wider()`
```{r}
#makeup a dataframe
analysis_result =
  tibble(
    group = c("treatment", "treatment", "placebo", "placebo"),
    time = c("pre", "post", "pre", "post"),
    group_mean = c(4,8,3,6)
  )
analysis_result

#extend the dataframe.
#let the names from variable "time" becomes new variables, and put the values from "group_mean" into the new columns
analysis_result %>%
  pivot_wider(
    names_from = "time",
    values_from = "group_mean"
  )
```


### binding the different tables: bind_rows()

1. first step: import each table
```{r}
fellowship_df = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "B3:D6") %>% #read the table
  mutate(movie = "fellowship_rings") #add a new column named "movie", with movie name "fellowship_rings"

two_towers_df = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "F3:H6") %>%
  mutate(movie = "two_towers")

return_df = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "J3:L6") %>%
  mutate(movie = "return_king")
```

2. second step: bind each rows together by using bind_rows() function
```{r}
lotr_df = 
  bind_rows(fellowship_df, two_towers_df, return_df) %>%
  janitor::clean_names() %>%
  relocate(movie) %>% #make "movie" to first column
  pivot_longer(
    female:male, #combine the variables(columns) from "female" to "male"
    names_to = "sex", #put the chosen variables' names into a new variable named "sex"
    values_to = "words" #put the values in the chosen variables into a new variable named "words"
  )
```

### Remeber: 
 - never use `rbinds()`, always use `bind_row()`.
 - never use `gather()`, always use `pivot_longer()`
 - never use `spread()`, always use `pivot_wider()`


## joining datasets
merge litters_df dataset into pups_df dataset:

1. read the datasets
```{r}
litters_df = 
  read_csv("./data/FAS_litters.csv") %>%
  janitor::clean_names() %>%
  relocate(litter_number)  %>% 
  separate(group, into = c("dose", "day_of_tx"), 3) %>% 
      #separate `group` variable into two different columns, named "dose", and "day_of_tx".
      # '3' :means spit the name after 3 characters (eg: con/7)
  mutate(dose = str_to_lower(dose))


 pups_df = 
  read_csv("./data/FAS_pups.csv") %>%
  janitor::clean_names() %>%
  mutate(sex = recode(sex, '1' = "male", '2' = "female")) #change `1` to `male`, `2` to `female`
        #'1' is a variable but not number, forcing R to recognize
   
```

2. Join the data: left_join()   
 - Since merge litters_df dataset into pups_df dataset, hence pups_df first, then litters_df
```{r}
fas_df = 
  left_join(pups_df, litters_df, by = "litter_number") %>%  #Join by "litter_number"
  arrange(litter_number) %>% 
  relocate(litter_number, dose, day_of_tx)
```

